A Quantitative Genetic Model of Background Selection in Humans

### 2023-09-08 manuscript/manuscript.tex
new updated ms from overleaf

\RequirePackage{cleveref}
\RequirePackage{xr}
\externaldocument{supplementary}

@@

% line numbers:
\RequirePackage{lineno}
%\modulolinenumbers[5]
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{black}}

@@

%\title{Towards a Complete Model of the Negative Selection Process in Humans}
%\title{An Evolutionary Quantitative Genetic Model of the Negative Selection Process in Humans}
\title{A Quantitative Genetic Linked Selection Model of Linked Selection in Humans}

\author{Vince Buffalo and Andrew Kern}

@@

\begin{abstract}

    Across the human genome, there are large-scale fluctuations in genetic
    diversity caused by the indirect effects of selection. This can be thought
    of as a ``linked selection signal" that reflects the impact of selection
    varying according to the placement of functional regions and recombination
    rates along the genome. Previous work has shown that negative selection
    against the steady influx of new deleterious mutations into conserved
    regions is the predominant mode of selection in humans. However, the
    theoretic model that underpins these results, classic Background Selection
    theory, is only applicable when new mutations are so deleterious that they
    cannot fix in the population. Here, we develop a statistical method based
    on a quantitative genetics view of the linked selection, which models the
    effects of weak draft created according to how polygenic additive fitness
    variance is distributed along the genome. We use a recent model that
    jointly predicts the equilibrium fitness variance and substitution rates
    due to both strong and weakly deleterious mutations, we estimate the
    distribution of fitness effects (DFE) and mutation rate across three human
    populations. While our model can accommodate weaker selection, we initially
    find evidence across three human populations of very strong selection
    against deleterious mutations consistent with previous work. However, the
    corollary predicted substitution rates for conserved regions are
    unreasonably low, and in disagreement with observed rates. We hypothesize
    this could be due to selected sites experiencing a further diminished
    population size due to selective interference. When we account for this in
    our method, we find evidence of weakly deleterious mutations in conserved
    regions which brings the predicted substitution rate into agreement with
    observations. However, these models lead to implausibly large mutation rate
    estimates. Overall, while our model of the genomic linked selection signal
    brings us a step towards uniting population and quantitative genetic
    selection models with the substitution process, our work suggests
    considerable uncertainty remains about the processes generating fitness
    variance in humans.

\end{abstract}

@@

The continual influx of new mutations into populations is the ultimate source
of all adaptations, but the vast majority of mutations either do not affect
fitness or are deleterious. Negative selection (i.e. purifying selection) works
to eliminate these deleterious mutations from the population; thus we expect
them to appear at low frequencies within populations
\parencite{Haldane1927-ga}, and be less likely to fix between lineages. The
best predictor of whether a new mutation will reduce fitness is if it occurs in
a functional region of the genome, which are often conserved over deep
phylogenetic timescales \parencite{Siepel2005-wh,Margulies2003-sn}. Since these
conserved regions reflect the product of millions of years of evolutionary
optimization, it follows that deleterious variants should be the overwhelming
majority of segregating variation in these regions. Furthermore, rare variation
in these regions is responsible for a significant portion of phenotypic
variation and disease in humans
\parencite{Zeng2018-ci,Lek2016-cb,Karczewski2020-ky,Tennessen2012-ge}.

Selection on both beneficial and deleterious variants perturbs the allele
frequencies of neighboring linked sites, a phenomenon known as linked selection
\parencite{Nordborg1996-nq,Maynard_Smith1974-zr,Barton1998-aj,Charlesworth1993-gb}.
Since deleterious variation is clustered in functional portions of the genome,
we expect linked selection to reduce levels of diversity around conserved
segments (e.g. genes, regulatory elements, etc.). The genomic arrangement of
these evolutionarily conserved regions and recombination rates create a
large-scale spatial signal of the linked selection process in genetic diversity
along the genome. Since genome-wide recombination maps and putatively conserved
segments are available for many species, there has been consistent effort to
fit models to this linked selection signal to estimate population genetic
parameters such as the strength of selection and the deleterious mutation rate
\parencite{Hudson1995-pt,McVicker2009-ax}, as well as distinguish and quantify
the roles of positive and negative selection and estimate the rate of
beneficial mutations \parencite{Elyashiv2016-vt,Murphy2022-sj}. In humans,
previous work has shown that negative selection plays the dominant role in
shaping megabase-scale patterns of diversity, with positive selection having a
nearly negligible impact \parencite{Murphy2022-sj}.

Prior work to model the reduction in linked diversity due to deleterious
mutations has largely relied on the classic Background Selection (BGS) model
\parencite{Charlesworth1993-gb,Nordborg1996-nq,Hudson1995-pt,Hudson1995-xc}.
While this model has been successful in explaining many patterns of diversity,
some its simplifying assumptions may distort inferences about the negative
selection process. First, since fixation probabilities ultimately depend on the
product of the deleterious selection coefficient ($s$) and population size
($N$), the efficacy of negative selection should depend on demography.
Unfortunately, accommodating demography into theoretic negative selection
models remains an open, difficult problem \parencite{Zeng2013-ep,Johri2020-oj}.
Second, building off classic models of mutation-selection balance
\parencite{Crow1970-wm,Kimura1966-bk}, the BGS model assumes that new mutations
are sufficiently deleterious that they are invariably driven to loss. Under
this assumption, the effect of selection is well-approximated by simply
rescaling the neutral coalescent by a reduction factor known as $B =
\nicefrac{N_e}{N}$ \parencite{Charlesworth2013-kl}. However, this simple
rescaling approach is not appropriate across parts of parameter space that are
relevant to natural populations \parencite{McVean2000-bt,Good2014-yz}. In
particular, the BGS model cannot accommodate the possibility of fixation for
mutations with $2Ns \ll 1$, leading diversity to be incorrectly predicted to
fall to zero as the strength of selection diminishes. This weak selection
problem is related to the wider problem of modeling the substitution rate of
deleterious mutation (i.e. the ``ratchet" rate) and incorporating the impact of
interference from other neighboring selected sites. Finally, the classic BGS
model assumes that selection occurring within other segments does not interfere
with selection occurring in other regions, i.e. no selective or
``Hill--Robertson" interference
\parencite{Hill1966-kd,McVean2000-bt,Felsenstein1974-xm}.

In this work, we use another class of linked selection models that derive from
quantitative genetics to address limitations of the classic BGS model. These
models express linked selection in terms of how polygenic fitness variance
creates weak genetic draft, and which increases pairwise coalescence rates.
While these models can theoretically accommodate additive fitness variance from
any source as long as its rate of change is not too rapid, we focus
specifically on a deleterious-mutations-only model of fitness variance from
\textcite{Santiago2016-mu}. This model is identical to BGS when selection
against deleterious mutations is strong, but it also correctly predicts the
reduction in diversity when selection is weak by jointly predicting the
deleterious substitution rate. We extend the Santiago and Caballero (hereafter
the SC16) model of the negative selection process so that it can be fit to
patterns of genome-wide diversity, according to the spatial distribution of
genomic features that could harbor deleterious fitness variation. Using forward
simulations, we show this model leads to more accurate estimates of the
distribution of fitness effects under weak selection. We apply our
composite-likelihood method to human population genomic data and provide new
estimates of the negative selection process in humans. However, we find that
there may be considerable uncertainty around these estimates if selective
interference is occurring across segments under purifying selection.

@@

% Similar to previous work, we aim to characterize selection processes through
% the signal left on nucleotide diversity at linked neutral sites. Although
% nucleotide diversity is a less informative statistic than the full site
% frequency spectrum, it is much easier to derive theoretic models of how linked
% selection alters expected pairwise coalescence times than to describe the
% entire genealogy. Still, inferences drawn from the linked selection signal
% about selective processes are only as accurate as the underlying theoretic
% model. Since we base our method on a fundamentally different theoretic approach
% to modeling linked selection, we briefly introduce the theory was extend to
% apply across whole-genomes.

Our work extends quantitative genetic models of linked selection
\parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Santiago2016-mu},
which model the effects linked selection in terms of polygenic additive fitness
variance ($V_A$). These linked selection models stem from
\textcite{Robertson1961-ho}, which in essence describes how heritable polygenic
fitness variation causes weak genetic draft. At the individual level Robertson
considered, selection generates autocorrelation as offspring from large
families tend to beget many descendents themselves (and likewise with small
families) when fitness is heritable. This same autocorrelation occurs at the
genomic level with linkage \parencite{Santiago1998-bs,Barton2000-zg}, as the
perturbations to a neutral allele's trajectory from its particular fitness
background tend to occur in the same direction across generations until the
background recombines off. These models quantify the total impact of the
autocorrelation generated by weak draft in terms of what we think of as a
\emph{draft-effective} population size $N_d$. The key insight is that in the
long-run, steady, weak draft created by the presence of additive genetic
fitness variance $V_A > 0$ acts like an extra variance in offspring number
would under pure drift \parencite{Wright1938-tv}. However, due to
autocorrelation, the cumulative effect of this fitness variance is inflated by
a factor of $Q^2$. Intuitively, the product $V_A Q^2$ represents the expected
total variance in reproductive success a neutral mutation experiences over its
lifetime in a system with weak selection draft.

Following \textcite{Robertson1961-ho} and \textcite{Santiago1995-hx}, we define
the draft-effective population size $N_d$ by including the total additional
variance created by heritable fitness into Wright's (\citeyear{Wright1938-tv})
equation for effective population size,

\begin{align}
    \label{eq:main_Ne}
    N_d = \frac{N}{Q^2 V_A + 1}
\end{align}
%
(c.f. \cite{Robertson1961-ho,Santiago1995-hx}; see Supplementary Materials Section
\ref{supp:theory} for a proof). The benefit of modeling draft with Robertson's
forward-time model is that the inflation factor is invariant with respect to
the particular fitness background the neutral allele becomes stochastically
associated with. By contrast, the difficulty with modeling draft backwards in
time is that the coalescence rates experienced by a lineage are not invariant
to which lineage was sampled due to its particular associated fitness
background.

Equation \eqref{eq:main_Ne} is general, since different modes of selection and
linkage can be accommodated by different expressions for $V_A$ and the
inflation factor $Q^2$ \parencite{Santiago1995-hx,Santiago1998-bs}. When
fitness variation has a multiplicative polygenic basis, as is often assumed for
genome-wide negative selection processes, the draft asymptotic effective
population size experienced by an arbitrary neutral site under the influence of
all $S$ linked regions is,

\begin{align}
    \label{eq:polygenic_Ne}
    N_d \approx N \exp\left(-\sum_{i=1}^S V_{A,i} \frac{Q_i^2}{2}\right)
\end{align}
%
where the factor of one-half comes from ignoring weak associations from
unlinked regions and chromosomes (see Supplementary Materials Section
\ref{supp:heritable-fitness}). In our genome-wide model, we consider the
summation in Equation \eqref{eq:polygenic_Ne} over non-overlapping segments $i
\in \{1, 2, \ldots, S\}$ each undergoing selection such that segment $i$
contributes additive fitness variance $V_{A,i}$ to the total additive genetic
fitness variance. The autocorrelation function for a neutral allele associated
with segment $i$ is $C(t) = [(1-r_i)(1-\kappa_i)]^t$, where $r_i$ is the
recombination fraction to the segment and $\kappa_i$ is the rate that the
associated fitness variance decays due to selective dynamics. Then, the
cumulative autocorrelation over the lifespan of the allele is,

\begin{align}
    \label{eq:Q}
    Q_i &= 1 + \sum_{j=1}^\infty \left[(1-r_i)(1-\kappa_i)\right]^j \nonumber \\
        &= \frac{1}{\kappa_i + r_i(1-\kappa_i)}.
\end{align}
%
(see Supplementary Materials Equation \ref{supp-eq:Qinf}). This general equation can accommodate
models of polygenic selection as long as the equilibrium additive fitness
variation $V_{A,i}$ can be specified and the change in variance due to
selection can be approximated as a geometric decay, i.e. $\Delta V_{A,i} =
-\kappa V_{A,i}$ \parencite{Bulmer1971-ae,Keightley1988-eq,Walsh2018-bt}. This
is usually a reasonable assumption since within-generation selection removes a
fraction of phenotypic variation from the population, and some fraction of that
is additive genetic variation \parencite{Bulmer1971-ae,Keightley1988-eq}.

The remaining required expressions are for the equilibrium additive fitness
variance $V_A$ and the decay rate in associated fitness $1-\kappa$. Fitness
variance could arise from positively or negatively selected alleles, but given
prior work has found negative selection against new deleterious mutations play
a dominant role in shaping genome-wide patterns of diversity
\parencite{McVicker2009-ax,Murphy2022-sj}, we focus on models of fitness
variance created by the continual influx of deleterious mutations into a
population. We imagine a mutation-selection process that creates fitness
variance through a continual influx of deleterious mutations occur at a rate of
$\mu$ per basepair per generation in a conserved region of $L$ basepairs, such
that the region-wide per generation diploid mutation rate is $U = 2 \mu L$.
Each mutation imposes a selective cost of $s$ in heterozygotes and $2s$ in
homozygotes, and fitness effects are multiplicative across sites.

Under this selection model, the additive genic fitness variance created by a
new mutation (at frequency $x=\nicefrac{1}{2N}$) is $2s^2 x(1-x) \approx
\nicefrac{s^2}{N}$. For the entire population of $2N$ chromosomes, the fresh
variance from mutation each generation in segment $i$ is $V_{m,i} \approx
U_is^2$ where $U_i = 2\mu L_i$ is the diploid mutation rate per generation
within the segment. Under the mutation-selection balance assumed by classic BGS
theory, an $L_i$-basepair segment has equilibrium genic variance $V_{a,i}^{BGS}
\approx U_i s$ (see Supplementary Materials Equation \ref{supp-eq:va_bgs}) and
thus $\kappa_i^{BGS} = s$. Substituting $V_{a,i}^{BGS}$ and $\kappa_i^{BGS}$ in
Equation \eqref{eq:polygenic_Ne} and simplifying, we have

\begin{align}
    N_d = N \exp \left( - \sum_i^S \frac{\mu L_i}{s(1 + r_i(1-s)/s)^2} \right) 
\end{align}
%
which is identical to the genome-wide model of background selection used in
previous studies \parencite{McVicker2009-ax,Elyashiv2016-vt,Murphy2022-sj}.
Thus, the classic background selection model is a special case of the more
general theory of Santiago and Caballero (\citeyear{Santiago2016-mu}), which
they had shown previously (\citeyear{Santiago1998-bs}).

However, when new mutations are only weakly deleterious, they can drift up in
frequency before their eventual loss or fixation. At this point, the number of
deleterious mutations per haplotype is no longer well-approximated by the
deterministic mutation-selection-balance theory; their dynamics are strongly
influenced by stochastic perturbations due to both drift and draft. In this
weak selection regime, classic BGS no longer accurately predicts levels of
linked diversity
\parencite{Charlesworth1993-gb,McVean2000-bt,Good2013-lp,Gordo2002-dr}.
Moreover, selection against weakly deleterious mutations alters the
genealogies, such that they are no longer well-approximated by a rescaled
neutral coalescent as assumed under the classic BGS model
\parencite{Przeworski1999-mb,OFallon2010-my,Higgs1995-xc}. To further
complicate matters, the distribution of the number of deleterious mutations
(and its corresponding fitness distribution) is no longer a stationary Poisson
distribution, instead becoming a traveling wave
\parencite{Rouzine2008-qz,Good2013-lp,Gessler1995-hz} towards reduced
population fitness (since we ignore beneficial mutations). When this occurs,
individuals in the least-loaded class are stochastically lost, fixing those
mutations with a click of ``Muller's ratchet"
\parencite{Muller1964-ki,Charlesworth1997-qn}. Unfortunately, determining the
rate of Muller's ratchet is another difficult, open problem
\parencite{Haigh1978-gt,Gordo2002-dr,Gessler1995-hz,Higgs1995-xc} related to
Hill--Robertson interference \parencite{Felsenstein1974-xm}.

Under qualitative genetic models of linked selection, these complications can
be avoided by finding a general expression for the equilibrium $V_A$ applicable
for both strong and weakly deleterious mutations. \textcite{Santiago2016-mu}
suggest that equilibrium fitness variance is lower than predicted by
$V_{a}^{BGS}$ once weakly deleterious mutations begin to have an appreciable
rate of fixation, $R > 0$, since each substitution (or click of Muller's
``ratchet") removes a segregating site and its heterozygosity becomes zero.
Thus, the steady-state genic variance under mutation and negative selection is
(omitting the segment index $i$ for clarity),

@@

where the condition $V_a \ge 0$ is met when the probability of fixation is less
than or equal to the neutral fixation probability of $\nicefrac{1}{2N}$, which
is held when considering deleterious mutations. This equation connects the
equilibrium additive genic fitness variance to the flux of new variation in
from new deleterious mutations, and the flux out due to their substitution (and
the decline in mean population fitness). When $R=0$, selection is so strong it
cannot fix, and the equilibrium fitness variation is due entirely to young rare
mutations before their extinction $V_a = V_a^{BGS} \approx Us$. Santiago and
Caballero derive Equation \eqref{eq:Va} through Fisher's Fundamental Theorem of
Natural Selection, but we find an alternative proof (\cite{Higgs1995-xc}; see
Supplementary Materials Section \ref{supp:weak-strong}). We also find that the
steady-state additive genic variance in Equation \eqref{eq:Va} results from
diffusion models with a flux of mutations into discrete sites
(\cite{Kimura1969-jw}).

%Our results differ from \parencite{Santiago2016-mu}, since we find from
%simulations their expression models the additive \emph{genic} fitness
%variation, not additive genetic fitness variation (see Section XXX). This is
%because Equation \eqref{eq:Va} does not capture the expected build-up of
%negative linkage disequilibria when selection is mildly deleterious.

>>>>

\RequirePackage{cleveref}
\RequirePackage{xr}

\newcommand*{\addFileDependency}[1]{
\typeout{(#1)}

\IfFileExists{#1}{}{\typeout{No file #1.}}
}\makeatother

\newcommand*{\myexternaldocument}[1]{%
\externaldocument{#1}%
\addFileDependency{#1.tex}%
\addFileDependency{#1.aux}%
}

\myexternaldocument{supplementary}

@@

% line numbers:
\RequirePackage[left, modulo]{lineno}
\modulolinenumbers[2]

% spacing
\RequirePackage{setspace}
\onehalfspacing

@@

%\title{Towards a Complete Model of the Negative Selection Process in Humans}
%\title{An Evolutionary Quantitative Genetic Model of the Negative Selection Process in Humans}
\title{A Quantitative Genetic Model of Background Selection in Humans}

\author[1,2,$\dag$]{Vince Buffalo}
\author[2]{Andrew D. Kern}
\affil[1]{\small{Department of Integrative Biology, University of California, Berkeley}}
\affil[2]{\small{Institute of Ecology and Evolution and Department of Biology, University of Oregon }}
\affil[*]{Corresponding author email: \href{vsb@berkeley.edu}{vsb@berkeley.edu}}

@@

\linenumbers
\begin{abstract}
    Across the human genome, there are large-scale fluctuations in genetic diversity caused by the indirect effects of selection. 
    This ``linked selection signal" reflects the impact of selection according to the physical placement of functional regions and recombination
    rates along chromosomes. Previous work has shown that purifying selection
    acting against the steady influx of new deleterious mutations at functional portions
    of the genome shapes patterns of genomic variation. To date, statistical
    efforts to estimate purifying selection parameters from linked selection models have relied on
    classic Background Selection theory, which is only applicable when new mutations are so deleterious that they cannot fix in the population. 
    Here, we develop a statistical method based
    on a quantitative genetics view of linked selection, that models how
    polygenic additive fitness variance distributed along the genome increases 
    the rate of stochastic allele frequency change. By 
    jointly predicting the equilibrium fitness variance and substitution rate
    due to both strong and weakly deleterious mutations, we estimate the
    distribution of fitness effects (DFE) and mutation rate across three geographically distinct human samples. While our model can accommodate weaker selection, we find evidence of strong selection operating similarly across all human samples. While our quantitative genetic model of linked selection fits better than previous models, substitution rates of the most constrained sites are poorly predicted. We find evidence that a model incorporating selective interference better predicts observed divergence, but overall our results ultimately suggest considerable uncertainty remains about the processes generating fitness variation in humans.
\end{abstract}

@@

The continual influx of new mutations into populations is the ultimate source of all adaptations, but the vast majority of mutations either do not affect fitness or are deleterious. Natural selection works to eliminate these deleterious mutations from the population, thus we expect them to appear at low frequencies within populations \parencite{Haldane1927-ga}, and be less likely to fix between lineages. Conserved genomic regions reflect the product of hundreds of millions of years of evolutionary optimization; thus the overwhelming majority of segregating variation in these regions will have deleterious fitness effects. Consequently, a good predictor of whether a new mutation will reduce fitness is if it occurs in a region of the genome that has been conserved over phylogenetic timescales \parencite{Siepel2005-wh,Margulies2003-sn}. Moreover, segregating rare variation in these regions is responsible for a significant proportion of the genetic contribution to phenotypic variation and disease in humans\parencite{Zeng2018-ci,Lek2016-cb,Karczewski2020-ky,Tennessen2012-ge}.

Selection on both beneficial and deleterious variants perturbs the allele frequencies of neighboring linked sites, a phenomenon known as linked selection
\parencite{Nordborg1996-nq,Maynard_Smith1974-zr,Barton1998-aj,Charlesworth1993-gb,Kaplan1989-sc}.
Since deleterious variation is clustered in functional portions of the genome,
we expect linked selection to reduce levels of diversity around evolutionarily constrained
segments (e.g. genes, regulatory elements, etc.). The genomic arrangement of
these evolutionarily conserved regions coupled with heterogeneous recombination rates create a
large-scale spatial signal of linked selection of genetic diversity along chromosomes. Since genome-wide recombination maps and functional annotations are available for many species, there has been consistent effort to fit models of linked selection to patterns of diversity. This general approach provides estimates of population genetic
parameters such as the strength of selection and the deleterious mutation rate
\parencite{Hudson1995-pt,McVicker2009-ax}, and potentially distinguishes the roles of positive and negative selection and estimate the rate of
beneficial mutations \parencite{Elyashiv2016-vt,Murphy2022-sj}. In humans,
previous work has shown that negative selection plays the dominant role in
shaping megabase-scale patterns of diversity, with positive selection having a
nearly negligible impact \parencite{Murphy2022-sj}.

Prior work to model the reduction in linked diversity due to deleterious
mutations has largely relied on the classic Background Selection (BGS) model
\parencite{Charlesworth1993-gb,Nordborg1996-nq,Hudson1995-pt,Hudson1995-xc}.
While the BGS model has been successful in fitting many patterns of diversity,
some of its simplifying assumptions may distort inferences about the selective process. First, since fixation probabilities ultimately depend on the
product of the deleterious selection coefficient ($s$) and population size
($N$), the efficacy of selection depends on past population sizes.
Unfortunately, accommodating such demography into purifying selection models remains an open, difficult problem \parencite{Zeng2013-ep,Johri2020-oj}.
Second as the BGS model builds off classic models of mutation-selection balance
\parencite{Crow1970-wm,Kimura1966-bk}, it assumes that new mutations are sufficiently deleterious that they are invariably driven to loss. Under
this assumption, the effect of selection is well-approximated by simply
rescaling the neutral coalescent by a reduction factor known as $B =
\nicefrac{N_e}{N}$ \parencite{Charlesworth2013-kl}. However, this simple
rescaling approach is not appropriate across parts of parameter space that are
relevant to natural populations \parencite{McVean2000-bt,Good2014-yz}. In
particular, the BGS model cannot accommodate the possibility of weakly deleterious mutations (those with fitness effects $2Ns \le 1$) reaching fixation, which leads to incorrect
predictions of diversity levels as the strength of selection diminishes. 
Finally, the classic BGS
model assumes that the selective dynamics occurring within a segment are not impacted by selection occurring outside the segment, i.e. no selective or
``Hill--Robertson" interference
\parencite{Hill1966-kd,McVean2000-bt,Felsenstein1974-xm}.

In this work, we use another class of linked selection models that derive from
quantitative genetics to address limitations of the classic BGS model \parencite{Santiago1998-bs, Santiago2016-mu,Santiago1995-hx,Robertson1961-ho}.
These models consider how polygenic fitness variance spread along the genome increases the variance of stochastic allele frequency change, as alleles become randomly linked to fitness backgrounds over time and their frequency trajectories are perturbed by selection at other sites. While these models can theoretically accommodate additive fitness variance from any source as long as its rate of change is not too rapid, we focus
specifically on a deleterious-mutations-only model of fitness variance from
\textcite{Santiago2016-mu}. This model is identical to BGS when selection
against deleterious mutations is strong, but it also correctly predicts the
reduction in diversity when selection is weak by jointly predicting the
deleterious substitution rate. We extend the Santiago and Caballero (hereafter
the SC16) model of the negative selection process so that it can be fit using a composite likelihood approach to patterns of genome-wide diversity, according to the spatial distribution of genomic features that could harbor deleterious fitness variation. Using forward simulations, we show this model leads to more accurate estimates of the distribution of fitness effects (DFE) under weak selection. We apply our
composite-likelihood method to human population genomic data and provide new
parameter estimates of the genome-wide impact of purifying selection in humans. 
We show that our new method is better able to predict the patterns of diversity along human chromosomes than previous models. 
However, our model leads to predictions of the deleterious substitution rate that disagree with observed levels of divergence. We discuss the potential causes and implications of such discrepancies and what it might mean for future efforts to fit linked selection models to genomic patterns of variation.

@@

Our work extends quantitative genetic models of linked selection
\parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Santiago2016-mu},
which model the effects of linked selection in terms of polygenic additive fitness
variance ($V_A$). Here we review the relevant theory before introducing our genome-wide extension. These linked selection models stem from
\textcite{Robertson1961-ho}, which in essence describes how polygenic additive
fitness variation increases the pace of stochastic allele frequency change, thus reducing effective population size. At the individual level Robertson considered, 
selection generates an autocorrelation in fecundity as offspring from large
families tend to beget many descendants themselves (and likewise with small
families) when fitness is heritable. This same across-generation autocorrelation occurs at the
genomic level due to linkage \parencite{Santiago1998-bs,Barton2000-zg}, as the
perturbations to a neutral allele's trajectory from its particular fitness
background tend to occur in the same direction across generations until the
background recombines off. Quantitative genetic models such as \textcite{Santiago1998-bs}
quantify the total impact of the autocorrelation generated by selection in terms of what we think of as a
\emph{fitness-effective} population size $N_f$ (to differentiate it from the \emph{drift}-effective population size, which is the size of the ideal population when there is no fitness variation).

The key insight is that in the long run, the steady presence of additive genetic fitness variance ($V_A > 0$) contributes an extra source of variance in offspring number beyond the variance expected under pure drift \parencite{Wright1938-tv}. However, because heritable fitness variation generates across-generation autocorrelation, the cumulative effect of this fitness variance on the variance in allele frequency change is inflated by a factor of $Q^2$. Intuitively, the product $V_A Q^2$ represents the expected
total variance in reproductive success a neutral mutation experiences over its lifetime in a system with weak selection at linked sites.

Following \textcite{Robertson1961-ho} and \textcite{Santiago1995-hx}, we define
the fitness-effective population size $N_f$ by including the total additional
variance created by heritable fitness into Wright's (\citeyear{Wright1938-tv})
equation for effective population size,

\begin{align}
    \label{eq:main_Ne}
    N_f = \frac{N}{Q^2 V_A + 1}
\end{align}
%
(c.f. \cite{Robertson1961-ho,Santiago1995-hx}; see Supplementary Materials Section
\ref{supp:theory} for a proof). The benefit of modeling linked selection with Robertson's
forward-time model is that the inflation factor is invariant with respect to the particular fitness background the neutral allele becomes stochastically associated with. By contrast,  modeling diversity levels under linked selection backwards in time requires tracking the particular associated fitness backgrounds, as coalescence rates experienced by a lineage are not invariant with respect to their fitness background.

Equation \eqref{eq:main_Ne} is general, since different modes of selection and
linkage can be accommodated by different expressions for $V_A$ and the
inflation factor $Q^2$ \parencite{Santiago1995-hx,Santiago1998-bs}. When
fitness variation has a multiplicative polygenic basis, as is often assumed for
genome-wide selection processes, the fitness-effective
population size experienced by an arbitrary neutral site under the influence of
all $S$ linked regions is,

\begin{align}
    \label{eq:polygenic_Ne}
    N_f \approx N \exp\left(-\sum_{i=1}^S V_{A,i} \frac{Q_i^2}{2}\right)
\end{align}
%
where the factor of one-half comes from ignoring the short-lived 
associations with the homologous chromosome, 
which have a weak effect on the focal allele (see Supplementary Materials Section
\ref{supp:heritable-fitness}). In our genome-wide model, we consider the
summation in Equation \eqref{eq:polygenic_Ne} over non-overlapping segments $i
\in \{1, 2, \ldots, S\}$ each undergoing selection such that segment $i$
contributes additive fitness variance $V_{A,i}$ to the total additive genetic
fitness variance. The impact this segment's fitness variance $V_{A,i}$ has on 
the fitness-effective population size is mediated by the autocorrelation term 
$Q_i$, which is a decaying function of the recombination rate between the 
segment and focal neutral allele. Specifically, the autocorrelation function
for a neutral allele associated with segment $i$ is $C(t) = [(1-r_i)(1-\kappa_i)]^t$, 
where $r_i$ is the recombination fraction to the segment and $\kappa_i$ is the rate that the
associated fitness variance decays due to selective dynamics. Then, the
cumulative autocorrelation over the lifespan of the allele is,

\begin{align}
    \label{eq:Q}
    Q_i &= 1 + \sum_{j=1}^\infty \left[(1-r_i)(1-\kappa_i)\right]^j \nonumber \\
        &= \frac{1}{\kappa_i + r_i(1-\kappa_i)}.
\end{align}
%
(see Supplementary Materials Equation \ref{supp-eq:Qinf}). This general equation can accommodate models of polygenic selection as long as the equilibrium additive fitness variation $V_{A,i}$ can be specified and the change in variance due to selection can be approximated as a geometric decay, i.e. $\Delta V_{A,i} = -\kappa V_{A,i}$ \parencite{Bulmer1971-ae,Keightley1988-eq,Walsh2018-bt, Santiago1998-bs}. This is usually a reasonable assumption since within-generation selection removes a fraction of phenotypic variation from the population, and some fraction of that is additive genetic variation \parencite{Bulmer1971-ae,Keightley1988-eq}.

The remaining required expressions are for the equilibrium additive fitness
variance $V_A$ and the decay rate in associated fitness $1-\kappa$. Fitness
variance could arise from beneficial or deleterious alleles, but given
prior work has found selection against new deleterious mutations in conserved regions plays a dominant role in shaping genome-wide patterns of diversity \parencite{McVicker2009-ax,Murphy2022-sj}, we focus specifically on purifying selection. We imagine a mutation-selection process that creates fitness variation as deleterious mutations enter a population at rate
$\mu$ per basepair per generation in a conserved region of $L$ basepairs, such
that the region-wide per generation diploid mutation rate is $U = 2 \mu L$.
Each mutation imposes a selective cost of $s$ in heterozygotes and $2s$ in
homozygotes, and fitness effects are multiplicative across sites.

Under this selection model, the additive genic fitness variance created by a
new mutation (at frequency $x=\nicefrac{1}{2N}$) is $2s^2 x(1-x) \approx
\nicefrac{s^2}{N}$. For the entire population of $2N$ chromosomes, the mutational variance input each generation in segment $i$ is $V_{M,i} \approx
U_is^2$ where $U_i = 2\mu L_i$ is the diploid mutation rate per generation
within the segment. Under the mutation-selection balance assumed by classic BGS
theory, an $L_i$-basepair segment has equilibrium genic variance $V_{a,i}^{BGS}
\approx U_i s$ (see Supplementary Materials Equation \ref{supp-eq:va_bgs}) and
thus $\kappa_i^{BGS} = s$. Substituting $V_{a,i}^{BGS}$ and $\kappa_i^{BGS}$ in
Equation \eqref{eq:polygenic_Ne} and simplifying, we have

\begin{align}
    N_f = N \exp \left( - \sum_i^S \frac{\mu L_i}{s(1 + r_i(1-s)/s)^2} \right) 
\end{align}
%
which is identical to the genome-wide model of background selection used in
previous studies \parencite{McVicker2009-ax,Elyashiv2016-vt,Murphy2022-sj}.
Thus, the classic background selection model is a special case of the more
general theory of Santiago and Caballero (\citeyear{Santiago2016-mu}), which
they had shown previously (\citeyear{Santiago1998-bs}).

However, when new mutations are only weakly deleterious, they can drift up in
frequency before their eventual loss or fixation. At this point, the number of
deleterious mutations per haplotype is no longer well-approximated by the
deterministic mutation-selection-balance theory, and their dynamics are strongly
influenced by stochastic perturbations due to both drift and linked selection. 
In this weak selection regime, classic BGS theory no longer accurately predicts levels of linked diversity
\parencite{Charlesworth1993-gb,McVean2000-bt,Good2013-lp,Gordo2002-dr}.
Moreover, selection against weakly deleterious mutations alters the topology of genealogies, such that they are no longer well-approximated by a rescaled
neutral coalescent as assumed under the classic BGS model
\parencite{Przeworski1999-mb,OFallon2010-my,Higgs1995-xc}. To further
complicate matters, the distribution of the number of deleterious mutations
(and its corresponding fitness distribution) is no longer a stationary Poisson distribution, instead becoming a traveling wave
\parencite{Rouzine2008-qz,Good2013-lp,Gessler1995-hz} towards increased numbers
of deleterious alleles per chromosome and reduced mean population fitness.
In asexual populations, substitutions correspond to a click of ``Muller's ratchet", which is 
the stochastic loss of the least-loaded class \parencite{Muller1964-ki,Charlesworth1997-qn}.
Unfortunately, determining the rate of Muller's ratchet is another difficult problem
\parencite{Haigh1978-gt,Gordo2002-dr,Gessler1995-hz,Higgs1995-xc} related to
Hill--Robertson interference \parencite{Felsenstein1974-xm}.

Under quantitative genetic models of linked selection, these complications can
be avoided by finding a general expression for the equilibrium $V_A$ applicable
to both strong and weakly deleterious mutations while concurrently modeling the rate
of fixation of deleterious alleles in the region, $R$. \textcite{Santiago2016-mu}
suggest that equilibrium fitness variance is lower than predicted by
$V_{a}^{BGS}$ once weakly deleterious mutations begin to have an appreciable
rate of fixation $R > 0$ per generation in the region. The substitution rate $R$
decreases fitness variance since each substitution removes a segregating site 
and thus its contribution to fitness variance. Thus the steady-state additive 
genetic variance of fitness under mutation and negative selection is
(omitting the segment index $i$ for clarity),

@@

where the condition $V_a \ge 0$ is met when the probability of fixation is less
than or equal to the neutral fixation probability of $\nicefrac{1}{2N}$, as is true for all deleterious mutations. This equation describes the
equilibrium additive genetic fitness variance as the balance of the flux of new variation in to the population from deleterious mutations, and the removal of variation due to their substitution (and the decline in mean population fitness). When $R=0$, selection is so strong deleterious alleles cannot fix, and the equilibrium fitness variation is due entirely to young rare mutations before their extinction $V_a = V_a^{BGS} \approx Us$. Santiago and
Caballero derive Equation \eqref{eq:Va} through Fisher's Fundamental Theorem of
Natural Selection, but we find an alternative proof (\cite{Higgs1995-xc}; see
Supplementary Materials Section \ref{supp:weak-strong}). We also find that the
steady-state additive genic variance in Equation \eqref{eq:Va} results from
diffusion models with a flux of mutations into discrete sites
(\cite{Kimura1969-jw}).
